---
title: "Assignment 2"
author: "Carmen Chen"
date: "10/4/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tibble)
library(ggplot2)
library(dplyr)
```


##Step 1: Data Wrangling

#Download the datasets
```{r get the data, cache=TRUE}
#individual dataset
ind <- "chs_individual.csv"
if (!file.exists(ind))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", destfile = ind)

ind <- read.csv(ind)
ind <- as.tibble(ind)

#regional dataset
reg <- "chs_regional.csv"
if (!file.exists(reg))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", destfile = reg)

reg <- read.csv(reg)
reg <- as.tibble(reg)
```

#Merge the data
```{r merge the data}
data <- merge(
  x = ind,
  y = reg,
  all.x = TRUE,
  all.y = TRUE,
  by.x = "townname",
  by.y = "townname"
)
```

#Check the dataset
```{r check for duplicates and missing values}
data %>% nrow()

colSums(is.na(data))

```
```{r data imputation}
data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(height = replace(height, is.na(height), mean(height, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(agepft = replace(agepft, is.na(agepft), mean(agepft, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(weight = replace(weight, is.na(weight), mean(weight, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(bmi = replace(bmi, is.na(bmi), mean(bmi, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(asthma  = replace(asthma, is.na(asthma), median(asthma, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(father_asthma = replace(father_asthma, is.na(father_asthma), median(father_asthma, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(mother_asthma = replace(mother_asthma, is.na(mother_asthma), median(mother_asthma, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(wheeze = replace(wheeze, is.na(wheeze), median(wheeze, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(hayfever = replace(hayfever, is.na(hayfever), median(hayfever, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(allergy = replace(allergy, is.na(allergy), median(allergy, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(educ_parent = replace(educ_parent, is.na(educ_parent), median(educ_parent, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(smoke = replace(smoke, is.na(smoke), median(smoke, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(gasstove = replace(gasstove, is.na(gasstove), median(gasstove, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(fev = replace(fev, is.na(fev), mean(fev, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(fvc = replace(fvc, is.na(fvc), mean(fvc, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(mmef = replace(mmef, is.na(mmef), mean(mmef, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(no_24hr = replace(no_24hr, is.na(no_24hr), mean(no_24hr, na.rm = TRUE)))

data <- data %>% 
     group_by(male, hispanic) %>%
     mutate(pm2_5_fr = replace(pm2_5_fr, is.na(pm2_5_fr), mean(pm2_5_fr, na.rm = TRUE)))

colSums(is.na(data))
```




#Creat categorical varialbes
```{r create obesity_level}
data$obesity_level <- as.factor(ifelse(data$bmi<14, "underweight",
                        ifelse(data$bmi<=22, "normal",
                               ifelse(data$bmi<=24, "overweight", "obese"))))

tapply(data$bmi, data$obesity_level, summary)
```

```{r create smoke_gas_exposure}
data$smoke_gas_exposure <- as.factor(
  ifelse(data$smoke==1 & data$gasstove==1, "Both",           
  ifelse(data$smoke==1 & data$gasstove==0, "Only second hand smoke",
  ifelse(data$smoke==0 & data$gasstove==0, "Neither", "Only gas stove"))))

table(data$smoke_gas_exposure, data$smoke)

table(data$smoke_gas_exposure, data$gasstove)

```


#Create summary tables

```{r summary tables}
library(data.table)

#by townname
data %>%
  group_by(townname,asthma) %>%
  summarise(mean_fev = mean(fev, na.rm = TRUE),
            sd_fev = sd(fev, na.rm =TRUE),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

#by sex
data %>%
  group_by(male,asthma) %>%
  summarise(mean_fev = mean(fev, na.rm = TRUE),
            sd_fev = sd(fev, na.rm =TRUE),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

#by obesity level
data %>%
  group_by(obesity_level,asthma) %>%
  summarise(mean_fev = mean(fev, na.rm = TRUE),
            sd_fev = sd(fev, na.rm =TRUE),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()

#by smoke_gas_exposure
data %>%
  group_by(smoke_gas_exposure,asthma) %>%
  summarise(mean_fev = mean(fev, na.rm = TRUE),
            sd_fev = sd(fev, na.rm =TRUE),
            n = n(),
            sd_asthma = sd(asthma, na.rm = TRUE)) %>%
  mutate(prop = prop.table(n)) %>%
  knitr::kable()
```



##Step 2: Looking at the Data (EDA)

1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r facet plot}
library(ggplot2)
ggplot(data) + 
  geom_point(mapping = aes(x = bmi, y = fev, color = townname)) + 
  geom_smooth(mapping = aes(x = bmi, y = fev, linetype = townname)) + 
  facet_wrap(~townname)

```

2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r histograms}
#by obesity level
ggplot(data, aes(x = fev)) +
  geom_histogram(aes(fill = obesity_level))

#by smoke/gas exposure
ggplot(data, aes(x = fev)) +
  geom_histogram(aes(fill = smoke_gas_exposure))
```


3. Barchart of BMI by smoke/gas exposure.
```{r barchart}
ggplot(data) +
  geom_bar(mapping = aes(x = obesity_level, fill = smoke_gas_exposure))
```

4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r boxplot}
#boxplot of FEV by obesity level
ggplot(data) +
  geom_boxplot(mapping = aes(x = obesity_level, y = fev))

#boxplot of FEV by smoke/gas exposure category
ggplot(data) +
  geom_boxplot(mapping = aes(x = smoke_gas_exposure, y = fev))
```

5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r leaflet, eval = knitr::is_html_output(excludes = "gfm")}
library(leaflet)

pm25.pal <- colorNumeric(c('darkgreen', 'goldenrod', 'brown'), domain=data$pm25_mass)

leaflet(data) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng = ~lon,
    label = ~paste0(round(pm25_mass, 2), '  PM2.5 mass'), color = ~ pm25.pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=pm25.pal, values=data$pm25_mass,
          title='PM2.5 mass', opacity=1)

```

6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r PM2.5 mass & FEV}
#general
ggplot(data, mapping = aes(x = pm25_mass, y = fev)) + 
  geom_point() + 
  geom_smooth()

#by sex
data$sex <- ifelse(data$male == 0, "female", "male")
ggplot(data, mapping = aes(x = pm25_mass, y = fev, color = sex, linetype = sex)) + 
  geom_point() + 
  geom_smooth()

```

