---
title: "Assignment 1"
Name: Carmen Chen
Date: 09/11/2021
output:
 - github_document:
   html_preview: false
 - html_document: default
always_allow_html: true
---

Steps
1. Given the formulated question from the assignment description, you will now conduct EDA Checklist items 2-4. First, download 2004 and 2019 data for all sites in California from the EPA Air Quality Data website. Read in the data using data.table(). For each of the two datasets, check the dimensions, headers, footers, variable names and variable types. Check for any data issues, particularly in the key variable we are analyzing. Make sure you write up a summary of all of your findings.

2. Combine the two years of data into one data frame. Use the Date variable to create a new column for year, which will serve as an identifier. Change the names of the key variables so that they are easier to refer to in your code.

3. Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). Summarize the spatial distribution of the monitoring sites.

4. Check for any missing or implausible values of PM in the combined dataset. Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.

5. Explore the main question of interest at three different spatial levels. Create exploratory plots (e.g. boxplots, histograms, line plots) and summary statistics that best suit each level of data. Be sure to write up explanations of what you observe in these data.
  state
  county
  site in Los Angeles
  

## Step 1: Read and check the data

```{r read in the data}
#Install only if we don't have the package
if(!require(data.table)){
  install.packages("data.table")
}

#load required package
library(data.table)

#Read in the data
data04 <- data.table::fread("Data04.csv")
data19 <- data.table::fread("Data19.csv")
```


For each of the two datasets, check the dimensions, headers, footers, variable names and variable types
```{r check dataset}
#2004
dim(data04)
head(data04)
tail(data04)

str(data04)

#2019
dim(data19)
head(data19)
tail(data19)

str(data19)
```


Check for any data issues, particularly in the key variable we are analyzing. Make sure you write up a summary of all of your findings.

```{r check data issues}
#2004
table(data04$Date)
table(data04$UNITS)
table(data04$STATE)

summary(data04$`Daily Mean PM2.5 Concentration`)
data04[`Daily Mean PM2.5 Concentration`<0][order(`Daily Mean PM2.5 Concentration`)]
data04rm <- data04[`Daily Mean PM2.5 Concentration`>=0]
mean(is.na(data04$`Daily Mean PM2.5 Concentration`))

summary(data04$SITE_LATITUDE)
summary(data04$SITE_LONGITUDE)


#2019
table(data19$Date)
table(data19$UNITS)
table(data19$STATE)

summary(data19$`Daily Mean PM2.5 Concentration`)
data19[`Daily Mean PM2.5 Concentration`<0][order(`Daily Mean PM2.5 Concentration`)]
data19rm <- data19[`Daily Mean PM2.5 Concentration`>=0]
mean(is.na(data19$`Daily Mean PM2.5 Concentration`))

summary(data19$SITE_LATITUDE)
summary(data19$SITE_LONGITUDE)
```

Summary: There is no missing value in the key varaibles we are analyzing. There are some negative values in daily mean PM2.5 concentration. The mean, median and maximum value of daily mean PM2.5 concentration in California decreased from 2004 to 2019. Throughout the 2004, the daily mean PM2.5 concentration in California has a mean value of 13.13 ug/m3 and median value of 10.10 ug/m3 with the range from 0 to 251 ug/m3. Throughout the 2019, the daily mean PM2.5 concentration in California has a mean value of 7.78 ug/m3 and median value of 6.5 ug/m3 with the range from 0 to 120.90 ug/m3.



#Step 2: combine and organize the data

```{r }
#Combine the two years of data into one data frame. 
total <- rbind(data04, data19)
totalrm <- rbind(data04rm, data19rm)

#Use the Date variable to create a new column for year, which will serve as an identifier. 
totalrm$Date <- as.POSIXct(totalrm$Date, format = "%m/%d/%Y")
totalrm$year <- format(totalrm$Date, format = "%Y")
table(totalrm$year)

#Change the names of the key variables so that they are easier to refer to in your code.
names(totalrm)[names(totalrm) == "Daily Mean PM2.5 Concentration"] <- "PM2.5"
names(totalrm)[names(totalrm) == "SITE_LATITUDE"] <- "lat"
names(totalrm)[names(totalrm) == "SITE_LONGITUDE"] <- "lon"
names(totalrm)[names(totalrm) == "Site Name"] <- "site"

```



#Step 3: Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). 

```{r leaflet, eval = knitr::is_html_output(excludes = "gfm")}
library(leaflet)

#Generating a color palette
year.pal <- colorFactor(c('darkgreen', 'goldenrod'), domain=totalrm$year)

#Map
leaflet(totalrm) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(lat=~lat, lng=~lon, color=~year.pal(year))
```


Summarize the spatial distribution of the monitoring sites.

Summary: The monitoring sites location scatter from North to South in California. They pretty much overlap with each other in 2004 and 2019. 


#Step 4: Check for any missing or implausible values of PM2.5 in the combined dataset. Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.

```{r check PM2.5}
mean(is.na(totalrm$PM2.5))
summary(total$PM2.5)
mean(total$PM2.5<0)
```
There is no missing value of PM2.5. However, there is implausible values less than 0 of PM2.5. The proportion of the implausible values is 0.39%.


#Step 5: Explore the main question of interest at three different spatial levels. Create exploratory plots (e.g. boxplots, histograms, line plots) and summary statistics that best suit each level of data. Be sure to write up explanations of what you observe in these data.

```{r}
library(ggplot2)
library(dplyr)
#state
ggplot(totalrm) +
  geom_boxplot(mapping = aes(x=year, y=PM2.5))

#county
ggplot(data=totalrm[totalrm$PM2.5<100]) + 
  geom_boxplot(mapping = aes(x=year, y=PM2.5))+
  facet_wrap(~COUNTY, scale="free")

#site in Los Angeles
ggplot(data=totalrm[totalrm$PM2.5<100 & totalrm$COUNTY=="Los Angeles"]) +
  geom_boxplot(mapping = aes(x=year, y=PM2.5)) +
  facet_wrap(~site, scale="free")

```

Overall, the PM2.5 level in 2004 is higher than in 2019 in state, county, and site levels. We can conclude that the PM2.5 decreased from 2004 to 2019. 
